Probing heavy ion collisions using quark and gluon jet substructure

### 2018-11-20 qg_ML.tex
one more run through text

Figure~\ref{fig:jetdistributons_pp_pbpb} shows the distributions of the aforementioned jet observables in pp and PbPb collisions for quark (blue shaded histograms) and gluon jets (orange shaded histograms). Both pp and PbPb jets include the grid discretization. Furthermore, jets in PbPb collisions are subtracted using the GridSub background subtraction method. Note that the jet mass and radial moments are infrared and collinear (IRC) safe observables with different weights on the particle angular distributions. In vacuum, they reflect the fact that gluon jets have larger jet mass and larger radial moments than quark jets due to the larger color charge and Casimir color factor $C_A>C_F$. In the \jw PbPb simulations, both quark and gluon jets have distributions that are modified towards larger values. On the other hand, we see a significant increase of the pixel multiplicity and decrease of the values of $p_T^D$ for quark and gluon jets, respectively, with the caveat that pixel multiplicity is unreliable in a heavy ion environment and the $p_T^D$ is an IRC unsafe observable.

@@

Such a study is however predicated on identifying collections of quark and gluon jets in both AA and pp collisions. Experiments that have a full hermetic coverage jet detectors can take advantage of kinematic and event selection constraints such as forward-rapidity vs mid-rapidity or photon/Z-jets vs dijets at varying \pt intervals in order to select these flavor specific collections. One could also use any of the aforementioned classification methods such as DCNNs or TD to select these jet collections and perform further differential analyses. Another intriguing possibility of such a model i.e that can pick out quenched jets vs unquenched jets, is to test it on high multiplicity pA or pp collisions. Measurements from the LHC~\cite{ALICE:2017jyt, Khachatryan:2016txc} have shown that these events are rather special in that they showcase AA like collective behavior but do not appear to have any significant jet quenching or parton energy loss signatures. This could be a test of QGP like behavior in these high multiplicity collisions which would in turn have wide reaching inferences regarding formation of the QGP in small systems.

>>>>

Figure~\ref{fig:jetdistributons_pp_pbpb} shows the distributions of the aforementioned jet observables in pp and PbPb collisions for quark (blue shaded histograms) and gluon jets (orange shaded histograms). Both pp and PbPb jets include the grid discretization. Furthermore, jets in PbPb collisions are subtracted using the GridSub background subtraction method. Note that the jet mass and radial moments are infrared and collinear (IRC) safe observables with different weights on the particle angular distributions. In vacuum, they reflect the fact that gluon jets have larger jet mass and larger radial moments than quark jets due to the larger color charge and Casimir color factor $C_A>C_F$. In the \jw PbPb simulations, both quark and gluon jets have distributions that are modified towards larger values. On the other hand, we see a significant increase of the pixel multiplicity and decrease of the values of $p_T^D$ for quark and gluon jets, respectively, with the caveat that pixel multiplicity is unreliable in a heavy ion environment and the $p_T^D$ is an IRC unsafe observable. Even though these are IRC unsafe, we study them for two primary purposes . Firstly, with our GridSub we are effectively combining these into discrete cells which reduces the unsafe behavior. Secondly these observables have been historically used in measurements of identifying quark and gluon jets in pp collisions at the LHC~\cite{ATL-PHYS-PUB-2017-017, CMS-DP-2016-070, CMS-PAS-JME-13-002}.

@@

Such a study is however predicated on identifying collections of quark and gluon jets in both AA and pp collisions. Experiments that have a full hermetic coverage jet detectors can take advantage of kinematic and event selection constraints such as forward-rapidity vs mid-rapidity or photon/Z-jets vs dijets at varying \pt~intervals in order to select these flavor specific collections. One could also use any of the aforementioned classification methods such as DCNNs or TD to select these jet collections and perform further differential analyses. Another intriguing possibility of such a model that can pick out quenched jets vs unquenched jets, is to test it on high multiplicity pA or pp collisions. Measurements from the LHC~\cite{ALICE:2017jyt, Khachatryan:2016txc} have shown that these events are rather special in that they showcase AA like collective behavior but do not appear to have any significant jet quenching or parton energy loss signatures. This could be a test of QGP like behavior in these high multiplicity collisions which would in turn have wide reaching inferences regarding formation of the QGP in small systems.
